---
title: "Chi Analysis"
author: "LoveA"
date: '2022-08-01'
output: html_document
---

                           Chi - Squared Analysis

Two types of Chi -Squared Test:

a) goodness of fit

b) Test of independence

Data: iris => from R data
Load libraries
```{r}
library(tidyverse)
library(patchwork)
```

To view the dataset 
```{r}
View(iris)
```

Converting numerical variables to categorical variable

```{r}
flower <- iris %>% 
  mutate(size = cut(Sepal.Length,
                    breaks = 3,
                    labels = c("Small", "Medium", "Large"))) %>% 
  select(Species, size)
```

Let's view our new dataset

```{r}
table(flower)
```

Chi Square Test of goodness of fit

Research Question: Is there statistically significant difference in the proportion
                   of flowers that are small, medium and large (alpha = 0.05)

* H0: The proportion of flowers that are small, medium sized and large are equal

* H1: The proportion of flower that are small, medium and large are not eqaul

```{r}
table(flower$size)
flower %>% 
  select(size) %>% 
  table() %>% 
  chisq.test()
```

The above line of code could also be written as
```{r}
chisq.test(table(flower$size))
```

p-value = 6.673e-07

Reject the H0(that the proportions are equal)
Because there is evidence that the probability of sample data providing a difference of this magnitude 
or more is < 0.05

Chi Squared Test of independencies

* H0: The variables are Independent
There is no relationship between the variables
knowing the value of one variable does not help to predict the value of the
other variable

* H1: The variables are dependent 
There is a relationship between the variables, knowing the value of one does
help to predict the value of the other

```{r}
chisq.test(table(flower))
```

```{r}
flower %>% 
  table %>% 
  chisq.test()
```

```{r}
table(flower) %>% 
  chisq.test()
```

p-value < 2.2e-16


